Canonical rings of $\bq$-divisors on projective space and Hirzebruch surfaces

### 2015-07-16 surfaces-write-up/surface-write-up.tex
add intro first two paragraphs

\section{Introduction}
For any Weil $\bq$-divisor

>>>>

\section{Introduction}
For any Weil $\bq$-divisor $D$ on a rational surface $X$, the graded
\textbf{canonical ring} $R(X, D) := \bigoplus_{d \geq 0} H^0(X, dD)$
has connections to rings of Hilbert and Siegel modular forms. In the
study of these canonical rings, it is natural to describe
presentations. We study the generators and relations of canonical
rings for projective spaces $\bp^m$ and Hirzebruch surfaces $\hirz_m$.
When $D$ is a general $\bq$-divisor on any $\bp^m$ or on any
$\hirz_m$, we give bounds on the generators and relations of
$R(X, D)$. In particular, we give a complete description of the
canonical ring when $X = \bp^m$ and $D$ is an $\bq$-divisor
containing one hypersurface and give stronger bounds when
$D$ is any effective $\bq$-divisor on $\bp^m$.

In part, this work is motivated by moving to generalize the work
of O'Dorney \cite{dorney:canonical} who gives similar descriptions
of canonical rings for general $\bq$-divisors on $\bp^1$. For
particular $\bq$ divisors on stacky curves, the work by
Voight--Zureick-Brown \cite{vzb:stacky} and Landesman--Ruhm--Zhang
\cite{lrz:spin-cring} on log spin canonical rings provides
useful bounds on rings of modular forms of Fuchsian groups.
The results of this work may have applications to Hilbert
and Siegel modular forms. \todo{reword last sentence}
